The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2019-04-27 rnw/Special Cases and Scenarios.Rnw
formatting changes for jss

\begin{small}
<<>>=
grisons$stage <- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
twophase(formula = tvol ~ stage, data = grisons, 
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  boundary_weights = "boundary_weights")
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
twophase(formula=tvol ~ mean + stddev + max + q75 + stage, data = grisons,
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  boundary_weights = "boundary_weights")
@
\end{small}

@@

\begin{small}
\begin{Schunk}
\begin{Sinput}
R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, 
+    data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster",  small_area = list(sa.col = "ismallold", areas = c("1"),
+    unbiased = TRUE))
Warning message:
At least one terrestrial cluster not entirely included within small area 1.
Zero mean residual assumption for small area maybe violated.
Check mean_Rc_x_hat_G and consider alternative estimator 'psmall' 
\end{Sinput}
\end{Schunk}
\end{small}

\begin{small}
<<>>=
psmall.clust <- twophase(formula = basal ~ stade + couver + melange, 
  data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  cluster = "cluster", small_area = list(sa.col = "ismallold", areas = c("1"),
  unbiased = TRUE), psmall = TRUE)
@
\end{small}

\begin{small}
<<>>=
extpsynth.clust$estimation
@
\end{small}

\begin{small}
<<>>=
psmall.clust$estimation
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
grisons[which(grisons$phase_id_3p == 2)[1], "mean"] <- NA
@
\end{small}

\begin{small}
\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning messages:
1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the first phase (s1) is missing
2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the zero phase (s0) is missing
3: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  1 rows deleted due to missingness in the set of auxiliary parameters for the 
  zero phase (s0) (1 terrestrial plots affected by deletion)
\end{Sinput}
\end{Schunk}
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
grisons[which(grisons$phase_id_3p == 2)[1], "q75"] <- NA
@
\end{small}

\begin{small}
\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning messages:
1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the first phase (s1) is missing
2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Changed the phase_id for 1 rows to the zero phase (s0) due to missingness in 
  the set of auxiliary parameters for the first phase (s1) (1 terrestrial 
  information no longer usable by this change)
\end{Sinput}
\end{Schunk}
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
grisons[which(grisons$phase_id_3p == 0)[1], "mean"] <- NA
@
\end{small}

\begin{small}
\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75,  data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning message:
In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  1 rows deleted due to missingness in the set of auxiliary parameters for the 
  zero phase (s0) (0 terrestrial plots affected by deletion)
\end{Sinput}
\end{Schunk}
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade == "300"), ]
@
\end{small}

\begin{small}
\begin{Schunk}
\begin{Sinput}
R> twophase(formula = basal ~ stade + couver + melange, data = zberg,
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster")
Error in check.mandatoryInputs(formula, data, phase_id) : 
 Level '300' of factor variable 'stade' existing in s1(s0)-but not in s2 sample.
 Calculation of coefficient not feasible.
\end{Sinput}
\end{Schunk}
\end{small}

>>>>

<<>>=
grisons$stage <- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
@

@@

<<eval=FALSE>>=
twophase(formula = tvol ~ stage, data = grisons, 
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  boundary_weights = "boundary_weights")
@

@@

<<eval=FALSE>>=
twophase(formula=tvol ~ mean + stddev + max + q75 + stage, data = grisons,
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  boundary_weights = "boundary_weights")
@

@@

\begin{Schunk}
\begin{Sinput}
R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, 
+    data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster",  small_area = list(sa.col = "ismallold", areas = c("1"),
+    unbiased = TRUE))
Warning message:
At least one terrestrial cluster not entirely included within small area 1.
Zero mean residual assumption for small area maybe violated.
Check mean_Rc_x_hat_G and consider alternative estimator 'psmall' 
\end{Sinput}
\end{Schunk}

<<>>=
psmall.clust <- twophase(formula = basal ~ stade + couver + melange, 
  data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  cluster = "cluster", small_area = list(sa.col = "ismallold", areas = c("1"),
  unbiased = TRUE), psmall = TRUE)
@

<<>>=
extpsynth.clust$estimation
@

<<>>=
psmall.clust$estimation
@

@@

<<eval=FALSE>>=
grisons[which(grisons$phase_id_3p == 2)[1], "mean"] <- NA
@

\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning messages:
1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the first phase (s1) is missing
2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the zero phase (s0) is missing
3: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  1 rows deleted due to missingness in the set of auxiliary parameters for the 
  zero phase (s0) (1 terrestrial plots affected by deletion)
\end{Sinput}
\end{Schunk}

@@

<<eval=FALSE>>=
grisons[which(grisons$phase_id_3p == 2)[1], "q75"] <- NA
@

\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning messages:
1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the first phase (s1) is missing
2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Changed the phase_id for 1 rows to the zero phase (s0) due to missingness in 
  the set of auxiliary parameters for the first phase (s1) (1 terrestrial 
  information no longer usable by this change)
\end{Sinput}
\end{Schunk}

@@

<<eval=FALSE>>=
grisons[which(grisons$phase_id_3p == 0)[1], "mean"] <- NA
@

\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75,  data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning message:
In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  1 rows deleted due to missingness in the set of auxiliary parameters for the 
  zero phase (s0) (0 terrestrial plots affected by deletion)
\end{Sinput}
\end{Schunk}

@@

<<eval=FALSE>>=
zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade == "300"), ]
@

\begin{Schunk}
\begin{Sinput}
R> twophase(formula = basal ~ stade + couver + melange, data = zberg,
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster")
Error in check.mandatoryInputs(formula, data, phase_id) : 
 Level '300' of factor variable 'stade' existing in s1(s0)-but not in s2 sample.
 Calculation of coefficient not feasible.
\end{Sinput}
\end{Schunk}

### 2019-04-27 rnw/Threephase_Estimators and their Application.Rnw
formatting changes for jss

\begin{small}
<<>>=
formula.rm <- tvol ~ mean
formula.fm <- tvol ~ mean + stddev + max + q75
@
\end{small}

@@

\begin{small}
<<>>=
reg3p_nex <- threephase(formula.s0 = formula.rm, formula.s1 = formula.fm, 
  data = grisons, phase_id = list(phase.col = "phase_id_3p", s1.id = 1,  
  terrgrid.id = 2), boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
summary(reg3p_nex)
@
\end{small}

@@

\begin{small}
<<>>=
truemeans.G <- data.frame(Intercept = rep(1, 4),
  mean = c(12.85, 12.21, 9.33, 10.45))
rownames(truemeans.G) <- c("A", "B", "C", "D")
@
\end{small}

@@

\begin{small}
<<>>=
extsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE),
  exhaustive = truemeans.G, boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
extsynth_3p$estimation
@
\end{small}

@@

\begin{small}
<<>>=
synth_3p <- threephase(formula.rm, formula.fm, data = grisons,
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = FALSE),
  exhaustive = truemeans.G, boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
synth_3p$estimation
@
\end{small}

@@

\begin{small}
<<>>=
extpsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE), 
  boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
extpsynth_3p$estimation
@
\end{small}

>>>>

<<>>=
formula.rm <- tvol ~ mean
formula.fm <- tvol ~ mean + stddev + max + q75
@

@@

<<>>=
reg3p_nex <- threephase(formula.s0 = formula.rm, formula.s1 = formula.fm, 
  data = grisons, phase_id = list(phase.col = "phase_id_3p", s1.id = 1,  
  terrgrid.id = 2), boundary_weights = "boundary_weights")
@

<<>>=
summary(reg3p_nex)
@

@@

<<>>=
truemeans.G <- data.frame(Intercept = rep(1, 4),
  mean = c(12.85, 12.21, 9.33, 10.45))
rownames(truemeans.G) <- c("A", "B", "C", "D")
@

@@

<<>>=
extsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE),
  exhaustive = truemeans.G, boundary_weights = "boundary_weights")
@

<<>>=
extsynth_3p$estimation
@

@@

<<>>=
synth_3p <- threephase(formula.rm, formula.fm, data = grisons,
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = FALSE),
  exhaustive = truemeans.G, boundary_weights = "boundary_weights")
@

<<>>=
synth_3p$estimation
@

@@

<<>>=
extpsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE), 
  boundary_weights = "boundary_weights")
@

<<>>=
extpsynth_3p$estimation
@

### 2019-04-27 rnw/Twophase_Estimators and their Application.Rnw
formatting changes for jss

In some national forest inventories it has been common to sample information over the entire country, i.e., also at places outside forest even including areas such as high mountain sides where no trees can actually grow. This is caused by the fact that in order to calculate totals by multiplying the estimated density (e.g., timber volume per hectare) with the area value of the inventory domain, the area of a country is well known, whereas the precise area of the forest is often much harder to determine. In \citet{mandallaz2014} it is demonstrated that by doing so, one usually inflates the sample by a large number of zeros for the target variable (i.e., at sample locations outside forest). It has been mathematically shown that by this procedure, the coefficient of determination (R$^2$) of the regression model can be overoptimistic as the zeros are easy to predict. This is particularly the case if the occurence of zeros correlate with an explanatory variable, such as the elevation above sea level. Furthermore, \citet{mandallaz2014} showed that this procedure is inefficient as it inflates the estimated variance. In order to mitigate this effect, a solution is to filter out most of the non-forested areas where one can certainly assume that no forest exists.

@@

\begin{small}
<<eval=FALSE>>=
library("forestinventory")
data("grisons", package = "forestinventory")
head(grisons)
@
\end{small}

\begin{small}
<<eval=TRUE, echo=FALSE>>=
library("forestinventory")
set.seed(666)
g.temp<- grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)]
 g.temp[,c(2,3,4,5,6,8)]<- round(g.temp[,c(2,3,4,5,6,8)], digits = 2)
print(head(g.temp, n = 10), row.names = seq(1:10))
rm(g.temp)
@
\end{small}

@@

\begin{small}
<<>>=
reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  boundary_weights = "boundary_weights")
@
\end{small}

@@

\begin{small}
<<>>=
summary(reg2p_nex)
@
\end{small}

@@

\begin{small}
<<eval=FALSE>>=
colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
  x = TRUE)$x)
@
\end{small}

@@

\begin{small}
<<>>=
true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)

reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
  data = grisons,
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  exhaustive = true.means.Z)
@
\end{small}

An alternative way to look at the estimation results without using the \code{summary()} is to query \code{reg2p_ex} directly:
\begin{small}
<<>>=
reg2p_ex$estimation
@
\end{small}
Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is essentially because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.

@@

In order to apply the pseudo small area estimator (PSMALL) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
\begin{small}
<<>>=
psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
  unbiased = TRUE), psmall = TRUE, boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
summary(psmall_2p)
@
\end{small}

The small area functions all return an \code{S3} object of class \code{"twophase"} with subclass \code{"smallarea"}. In addition to global estimation, the \code{estimation} object now comprises the estimates and variances for all small areas (column \code{area}). We can view the sample sizes by looking into the object itself
\begin{small}
<<>>=
psmall_2p$samplesizes
@
\end{small}

@@

\begin{small}
<<>>=
extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
  unbiased = TRUE), boundary_weights = "boundary_weights")
extpsynth_2p$estimation
@
\end{small}

@@

\begin{small}
<<>>=
psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
  unbiased = FALSE), boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
psynth_2p$estimation
@
\end{small}

@@

For the \code{grisons} data set, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by
\begin{small}
<<eval=FALSE>>=
colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
  x = TRUE)$x)
@
\end{small}
\begin{small}
<<>>=
true.means.Z.G <- data.frame(Intercept = rep(1, 4),
  mean = c(12.85, 12.21, 9.33, 10.45),
  stddev = c(9.31, 9.47, 7.90, 8.36),
  max = c(34.92, 35.36, 28.81, 30.22),
  q75 = c(19.77, 19.16, 15.40, 16.91))
rownames(true.means.Z.G) <- c("A", "B", "C", "D")
@
\end{small}
\begin{small}
<<>>=
true.means.Z.G
@
\end{small}

The extended synthetic estimator (EXTSYNTH) can then be applied by
\begin{small}
<<>>=
extsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col ="smallarea", areas = c("A", "B"),
  unbiased = TRUE), exhaustive = true.means.Z.G)
@
\end{small}

\begin{small}
<<>>=
extsynth_2p$estimation
@
\end{small}

>>>>

In some national forest inventories it has been common to sample information over the entire country, i.e., also at places outside forest even including areas such as high mountain sides where no trees can actually grow. This is caused by the fact that in order to calculate totals by multiplying the estimated density (e.g., timber volume per hectare) with the area value of the inventory domain, the area of a country is well known, whereas the precise area of the forest is often much harder to determine. In \citet{mandallaz2014} it is demonstrated that by doing so, one usually inflates the sample by a large number of zeros for the target variable (i.e., at sample locations outside forest). It has been mathematically shown that by this procedure, the coefficient of determination (R$^2$) of the regression model can be overoptimistic as the zeros are easy to predict. This is particularly the case if the occurence of zeros correlates with an explanatory variable, such as the elevation above sea level. Furthermore, \citet{mandallaz2014} showed that this procedure is inefficient as it inflates the estimated variance. In order to mitigate this effect, a solution is to filter out most of the non-forested areas where one can certainly assume that no forest exists.

@@

<<eval=FALSE>>=
library("forestinventory")
data("grisons", package = "forestinventory")
head(grisons)
@

<<eval=TRUE, echo=FALSE>>=
library("forestinventory")
set.seed(666)
g.temp<- grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)]
 g.temp[,c(2,3,4,5,6,8)]<- round(g.temp[,c(2,3,4,5,6,8)], digits = 2)
print(head(g.temp, n = 10), row.names = seq(1:10))
rm(g.temp)
@

@@

<<>>=
reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  boundary_weights = "boundary_weights")
@

@@

<<>>=
summary(reg2p_nex)
@

@@

<<eval=FALSE>>=
colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
  x = TRUE)$x)
@

@@

<<>>=
true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)

reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
  data = grisons,
  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  exhaustive = true.means.Z)
@

An alternative way to look at the estimation results without using the \code{summary()} is to query \code{reg2p_ex} directly:

<<>>=
reg2p_ex$estimation
@

Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is essentially because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.

@@

In order to apply the pseudo small area estimator (PSMALL) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:

<<>>=
psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
  unbiased = TRUE), psmall = TRUE, boundary_weights = "boundary_weights")
@

<<>>=
summary(psmall_2p)
@

The small area functions all return an \code{S3} object of class \code{"twophase"} with subclass \code{"smallarea"}. In addition to global estimation, the \code{estimation} object now comprises the estimates and variances for all small areas (column \code{area}). We can view the sample sizes by looking into the object itself

<<>>=
psmall_2p$samplesizes
@

@@

<<>>=
extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
  unbiased = TRUE), boundary_weights = "boundary_weights")
extpsynth_2p$estimation
@

@@

<<>>=
psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
  unbiased = FALSE), boundary_weights = "boundary_weights")
@

<<>>=
psynth_2p$estimation
@

@@

For the \code{grisons} data set, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by

<<eval=FALSE>>=
colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
  x = TRUE)$x)
@

<<>>=
true.means.Z.G <- data.frame(Intercept = rep(1, 4),
  mean = c(12.85, 12.21, 9.33, 10.45),
  stddev = c(9.31, 9.47, 7.90, 8.36),
  max = c(34.92, 35.36, 28.81, 30.22),
  q75 = c(19.77, 19.16, 15.40, 16.91))
rownames(true.means.Z.G) <- c("A", "B", "C", "D")
@

<<>>=
true.means.Z.G
@

The extended synthetic estimator (EXTSYNTH) can then be applied by

<<>>=
extsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
  small_area = list(sa.col ="smallarea", areas = c("A", "B"),
  unbiased = TRUE), exhaustive = true.means.Z.G)
@

<<>>=
extsynth_2p$estimation
@
